### 2016-06-14 thesis/methods/methods.tex
tmp commit, added .swp in .gitignore

\paragraph{\textit{Raw data normalisation}}

@@

\paragraph{\textit{ Raw data standardisation }}

The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.

\paragraph{\textit{ Raw data standardisation }}

@@

\paragraph{\textit{ Raw data standardisation }}

The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.

@@

@@

\subsubsection{\textbf{Reformatting the raw data}}
\label{subsec:icgcreformat}

%TODO: add in how I processed the ICGC data format into genes by samples.
TODO: add in how I processed the ICGC data format into genes by samples.

%TODO: change the title name?
\subsubsection{\textbf{\gls{icgc} sample exclusion criteria}}
\label{subsec:icgcsampleexclusion}

\paragraph{\textit{Sample BMI availability}}
Some of the samples did not have either height or weight data, or both.
These samples were removed from the analyses.

\paragraph{\textit{Sample ID conversion and matching}}

%TODO: check the actual ID stripping process
%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample

\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID from the \gls{rnaseq} data matched with the \gls{tcga} ID in the clinical data.
After stripping the ID, samples were checked for any overlaps.
Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
Any sample that did not have either clinical or \gls{rnaseq} data were removed, so both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.

\begin{table}[h]
	\caption{Summary of the total number of samples included in the analyses for each cancer type}
	\label{tab:samplesize}
	\begin{center}
		\begin{tabular}{cc}
			\textbf{Cancer Type}   & \textbf{Total} \\
			\hline
			\rule{0pt}{2.25ex}BLCA & 261   \\
			CESC                   & 224   \\
			COAD                   & 226   \\
			KIRP                   & 124   \\
			LIHC                   & 264   \\
			READ                   & 73    \\
			SKCM                   & 218   \\
			UCEC                   & 482   \\
		\end{tabular}
	\end{center}
\end{table}

\subsection{Normalisation of raw Creighton data}
\label{subsec:normcrdata}

\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
\label{subsec:crobsgene}

The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

>>>>

\subsubsection{\textbf{Raw data normalisation}}

@@

\subsubsection{\textbf{Raw data standardisation}}

The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.

\subsubsection{\textbf{Gene probe ID conversion}}

Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
Likewise, the obesity associated gene probes were also converted into gene symbols where required.

\subsubsection{\textbf{Sample \gls{bmi} calculation}}

\gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.

@@

\subsubsection{\textbf{Raw data standardisation}}

The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.

\subsubsection{\textbf{Gene probe ID conversion}}

Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

@@

\subsubsection{\textbf{Raw data normalisation}}

The raw \gls{rnaseq} data for each cancer types from \cref{subsec:tcga} had 1 added, and then logged to the power of 10.

\subsubsection{\textbf{Raw data standardisation}}

The normalised data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.

@@

\subsubsection{\textbf{Raw data format conversion}}
%TODO: add in how I processed the ICGC data format into genes by samples.
The raw \gls{rnaseq} data for each cancer types were formatted so that the rows corresponded to the genes and the columns to the samples.
% TODO: find out what functions I used to reformat the icgc data
This was done using the \textit{dplyr} package in \gls{r}.

\subsubsection{\textbf{Sample ID conversion}}
%TODO: add details on how I stripped the ICGC ID into TCGA ID
%TODO: check the actual ID stripping process
%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample
\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID matched with the \gls{tcga} ID in the clinical data.
After the IDs were stripped, samples were checked for any overlaps.
Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

\subsubsection{\textbf{\gls{icgc} sample exclusion criteria}}
\label{subsec:icgcsampleexclusion}

% \subsubsection{\textbf{Sample BMI availability}}

Some of the samples did not have either height or weight data, or both.
These samples were removed from the analyses.
In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
Any sample that did not have either clinical or \gls{rnaseq} data were removed.
This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.

%TODO: May need more details on the sample (bmi status, etc.)
\begin{table}[h]
	\caption{Summary of the total number of samples included in the analyses for each cancer type}
	\label{tab:samplesize}
	\begin{center}
		\begin{tabular}{cc}
			\textbf{Cancer Type}   & \textbf{Total} \\
			\hline
			\rule{0pt}{2.25ex}BLCA & 261   \\
			CESC                   & 224   \\
			COAD                   & 226   \\
			KIRP                   & 124   \\
			LIHC                   & 264   \\
			READ                   & 73    \\
			SKCM                   & 218   \\
			UCEC                   & 482   \\
		\end{tabular}
	\end{center}
\end{table}

\section{Plot creation}
\label{sec:plot_creation}

\subsection{Heatmaps}
\label{sub:heatmaps}

\subsection{Box plots}
\label{sub:box_plots}

\subsection{Scatter plots}
\label{sub:scatter_plots}

\section{Gene expression analysis}
\label{sec:gene_expression_analysis}

\section{Metagene analysis}
\label{sec:metagene_analysis}

\section{Pathway enrichment analysis}
\label{sec:pathway_enrichment_analysis}
